Competing Bandits:\\ The Perils of Exploration under Competition \thanks{This is a merged and final version of two conference papers, \citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19}, with a unified and streamlined presentation.}

### 2020-07-13 journal paper/content/sec-intro.tex
Reworded in "network effects".

\ascomment{The xhdr below is basically from Guy, with minor edits by Alex.}

\xhdr{Economic interpretation: network effects of data.}
Our model speaks to the ongoing debate on the role that data plays in the digital economy, \gaedit{which has been at the forefront of policy discussions on how to regulate major, data-intensive, digital platforms \citep{competition2019digital, scott2019committee}.} One fundamental question in this debate is whether data can serve a similar role as traditional ``network effects", \asedit{creating scenarios when only one firm can function in the market \citep{Rysman09, jullien2019economics}.}
%whereby, when these effects are present, in many cases only one firm can function in the market, leading to competition \emph{for} the market being more important than competition \emph{in} the market \citep{Rysman09, jullien2019economics}.
\asedit{The death spiral/amplification effects mentioned above have a similar flavor as network effects: a relatively small amount of exploration (resp., data advantage)  gets amplified under competition and causes the firm to be starved out of users (resp., take over most of the market).}
%\gaedit{We further find that a small data advantage for one firm gets amplified under competition and leads to that firm taking the entire market, showing that data can provide a similar incumbency advantage as those provided by traditional network effects and can serve as a barrier to entry in online markets.}
A distinctive feature of our approach is that we explicitly model the learning problem of the firms and consider them deploying algorithms for solving this problem.  Thus, we do not explicitly model the network effects, but they arise endogenously from our setup.

@@

\ascomment{Guy: citing \citet{varian2018artificial} might serve us, even if it is a more informal discussion. The existence of the policy discussion is worth mentioning, too (with citations).}

>>>>

\ascomment{The xhdr below is basically from Guy, with some rewordings by Alex.}

\xhdr{Economic interpretation: network effects of data.}
\asedit{Our model speaks to policy discussions on regulating data-intensive digital platforms \citep{competition2019digital, scott2019committee}, and particularly to the ongoing debate on the role of data in the digital economy.} One fundamental question in this debate is whether data can serve a similar role as traditional ``network effects", \asedit{creating scenarios when only one firm can function in the market \citep{Rysman09, jullien2019economics}.}
%whereby, when these effects are present, in many cases only one firm can function in the market, leading to competition \emph{for} the market being more important than competition \emph{in} the market \citep{Rysman09, jullien2019economics}.
\asedit{The death spiral/amplification effects mentioned above have a similar flavor as network effects: a relatively small amount of exploration (resp., data advantage)  gets amplified under competition and causes the firm to be starved out of users (resp., take over most of the market).}
%\gaedit{We further find that a small data advantage for one firm gets amplified under competition and leads to that firm taking the entire market, showing that data can provide a similar incumbency advantage as those provided by traditional network effects and can serve as a barrier to entry in online markets.}
A distinctive feature of our approach is that we explicitly model the learning problem of the firms and consider them deploying algorithms for solving this problem.  Thus, we do not explicitly model the network effects, but they arise endogenously from our setup.

@@
